---
title: "Assignment 6 - SVMs and Unsupervised Learning"
author: "Slava Mikhaylov"
output: html_document
---

Assignments for the course focus on practical aspects of the concepts covered in the lectures. Assignments are based on the material covered in James et al. You will start working on the assignment in the lab sessions after the lectures, but may need to finish them after class.

We open solutions on Wednesdays after 9pm, you will have until then to work on your assignments and submit them via Moodle. 


### Exercise 6.1

In this problem, you will use support vector approaches in order to predict whether a given car gets high or low gas mileage based on the `Auto` data set.

(a) Create a binary variable that takes on a 1 for cars with gas mileage above the median, and a 0 for cars with gas mileage below the median.
(b) Fit a support vector classifier to the data with various values of cost, in order to predict whether a car gets high or low gas mileage. Report the cross-validation errors associated with different values of this parameter. Comment on your results.
(c) Now repeat (b), this time using SVMs with radial and polynomial basis kernels, with different values of gamma and degree and cost. Comment on your results.
(d) Make some plots to back up your assertions in (b) and (c). 
*Hint: In the lab, we used the `plot()` function for `svm` objects only in cases with $p = 2$. When $p > 2$, you can use the `plot()` function to create plots displaying pairs of variables at a time. Essentially, instead of typing*
```{r eval=FALSE}
plot(svmfit , dat)
```
*where `svmfit` contains your fitted model and `dat` is a data frame containing your data, you can type*
```{r eval=FALSE}
plot(svmfit, dat, x1 ~ x4)
```
*in order to plot just the first and fourth variables. However, you must replace $x1$ and $x4$ with the correct variable names. To find out more, type `?plot.svm`.*


### Exercise 6.2

In this problem, you will generate simulated data, and then perform PCA and K-means clustering on the data.

(a) Generate a simulated data set with 20 observations in each of three classes (i.e. 60 observations total), and 50 variables.
*Hint: There are a number of functions in `R` that you can use to generate data. One example is the `rnorm()` function; `runif()` is another option. Be sure to add a mean shift to the observations in each class so that there are three distinct classes.*
(b) Perform PCA on the 60 observations and plot the first two principal component score vectors. Use a different color to indicate the observations in each of the three classes. If the three classes appear separated in this plot, then continue on to part (c). If not, then return to part (a) and modify the simulation so that there is greater separation between the three classes. Do not continue to part (c) until the three classes show at least some separation in the first two principal component score vectors.
(c) Perform $K$-means clustering of the observations with $K = 3$. How well do the clusters that you obtained in $K$-means clustering compare to the true class labels?
*Hint: You can use the `table()` function in `R` to compare the true class labels to the class labels obtained by clustering. Be careful how you interpret the results: $K$-means clustering will arbitrarily number the clusters, so you cannot simply check whether the true class labels and clustering labels are the same.*
(d) Perform $K$-means clustering with $K = 2$. Describe your results.
(e) Now perform $K$-means clustering with $K = 4$, and describe your
results.
(f) Now perform $K$-means clustering with $K = 3$ on the first two principal component score vectors, rather than on the raw data. That is, perform $K$-means clustering on the $60 \times 2$ matrix of which the first column is the first principal component score vector, and the second column is the second principal component score vector. Comment on the results.
(g) Using the `scale()` function, perform $K$-means clustering with $K = 3$ on the data after scaling each variable to have standard deviation one. How do these results compare to those obtained in (b)? Explain.


### Exercise 6.3 (Optional)

This problem involves the `OJ` data set which is part of the `ISLR` package.

(a) Create a training set containing a random sample of 800 observations, and a test set containing the remaining observations.
(b) Fit a support vector classifier to the training data using `cost=0.01`, with `Purchase` as the response and the other variables as predictors. Use the `summary()` function to produce summary statistics, and describe the results obtained.
(c) What are the training and test error rates?
(d) Use the `tune()` function to select an optimal cost. Consider values in the range 0.01 to 10.
(e) Compute the training and test error rates using this new value for `cost`.
(f) Repeat parts (b) through (e) using a support vector machine with a radial kernel. Use the default value for `gamma`.
(g) Repeat parts (b) through (e) using a support vector machine with a polynomial kernel. Set `degree=2`.
(h) Overall, which approach seems to give the best results on this data?